---
title: "Lab 1 - Data visualization"
format: html
editor: visual
author: Aiden Mansfield
---

# Questions

```{r}
install.packages("tidyverse")
library(tidyverse)
```

## Part 1

### Question 1

```{r}
ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 100) +
  labs(
    x = "Population density",
    y = "Count",
    title = "Population density of midwesten counties",
    subtitle = "Binwidth = 100"
  )

ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 1000) +
  labs(
    x = "Population density",
    y = "Count",
    title = "Population density of midwesten counties",
    subtitle = "Binwidth = 1000"
  )

ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 10000) +
  labs(
    x = "Population density",
    y = "Count",
    title = "Population density of midwesten counties",
    subtitle = "Binwidth = 10000"
  )

ggplot(midwest, aes(x = popdensity)) +
  geom_histogram(binwidth = 100000) +
  labs(
    x = "Population density",
    y = "Count",
    title = "Population density of midwesten counties",
    subtitle = "Binwidth = 100000"
  )
```

A binwdith of 1000 is likley the most appropriate. It balances smoothness and detail while still showing that the data is right tailed.

### Question 2

```{r}
ggplot(midwest, aes(x = popdensity)) +
  geom_boxplot() +
  labs(
    x = "Population density",
    y = "Count",
    title = "Population density of midwesten counties",
  )
```

The data is strongly skewed to the right. One outlier is Cook County, IL because it is significantly larger than the typical county, located on the far right of the x axis as an isolated point.

### Question 3

```{r}
ggplot(
  data = midwest,
  mapping =aes(x = percollege, y = percbelowpoverty, color = state, shape = state)
) +  geom_point() +
  labs(
    x = "Percentage of people with college degree",
    y = "Percentage of people below poverty",
    title = "College Education vs Poverty in Midwestern Counties",
    color = "State",
    shape = "State"
 )
```

A negative relationship between college education and poverty exists. Generally speaking, as the percentage of people with a collage degree increases, the percentage below poverty generally decreases. One outlier is Jackson, IL with 36.6% percollege and 28.4% perbelowpoverty which is higher than other counties with similar education levels.

### Question 4

```{r}
ggplot(
  data = midwest,
  mapping =aes(x = percollege, y = percbelowpoverty, color = state, shape = state)
) +  geom_point() + geom_smooth(se = FALSE) + facet_wrap(~state) + 
  labs(
    x = "Percentage of people with college degree",
    y = "Percentage of people below poverty",
    title = "College Education vs Poverty in Midwestern Counties",
    color = "State",
    shape = "State"
 )
```

I prefer this plot because it's easier to interpret the data by providing clearer breakdowns by state.

### Question 5

```{r}
ggplot(
  data = midwest,
  mapping =aes(x = state, y = area, color = state, shape = state)
) +  geom_boxplot() + geom_smooth(se = FALSE) + facet_wrap(~state) + 
  labs(
    x = "State",
    y = "County Area",
    title = "Distribution of County Area by State",
)
```

Typical county sizes are largest in Michigan and smallest in Ohio. Illinois, Indiana, and Wisconsin are somewhere in between. Variability in county size is greatest in Michigan and smallest in Ohio. Wisconsin shows more spread than Illinois or Indiana. The largest county is Marquette in Michigan with an area of 0.110.

### Question 6

### Question 7

## Part 2

**Enough about the Midwest!**

### Question 8

### Question 9
